---
title: "DDA2081_report"
author: "Jingyi Zong"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Paper report

## download data

Many of the databases in this article come from the GEO website. We use some packages in R to directly import the data into R, without the need to download it. Below is our code for downloading the GEO database (taking database 3 as an example).

```{r download data}
#BiocManager::install('GEOquery')
#BiocManager::install('limma',force = TRUE)
#BiocManager::install('umap',force = TRUE)
#idprob = AnnoProbe::idmap("GPL570",type = 'soft') # Find the gene information corresponding to the probe ID
#library(GEOquery)
#library(limma)
#library(umap)
#gset <- getGEO("GSE39582", GSEMatrix =TRUE, getGPL=FALSE)
#if (length(gset) > 1) idx <- grep("GPL570", attr(gset, "names")) else idx <- 1
#gset <- gset[[idx]]
#ex <- exprs(gset)
pd <- pData(gset)
```

## How to obtain the four genes needed for research

### DimensionReduction_PNAS.code

Import the files:"Tcolon_new.csv","Ncolon_new.csv"

```{r translate MATLAB code into R and explain the code}
T <- as.data.frame((read.csv('Tcolon_new.csv')))# tumor samples
N <- as.data.frame((read.csv('Ncolon_new.csv')))# normal samples
```

Identify the genes with the greatest differences between tumor people and normal people among 1991 genes.

```{r translate MATLAB code into R and explain the code}
m <- ncol(T) # the number of tumor samples
n <- ncol(N) # the number of normal samples

Cm <- Cv <- Cs <- numeric(nrow(T))

# Calculate means, variances and coefficient of variation
for (i in 1:nrow(T)) {
  tm <- mean(as.numeric(T[i,]))
  nm <- mean(as.numeric(N[i,]))
  tv <- sd(as.numeric(T[i,]))
  nv <- sd(as.numeric(N[i,]))
  ts <- tv/tm
  ns <- nv/nm
  Cm[i] <- abs((tm-nm)/nm)
  Cv[i] <- abs((tv-nv)/nv)
  Cs[i] <- abs((ts-ns)/ns)
}

# Sort the metrics and retrieve indices
Icm <- order(Cm)
Icv <- order(Cv)
Ics <- order(Cs)

#Jm <- sort(Icm[(length(Icm)-18):length(Icm)]) # % 1%
Jm <- sort(Icm[1:19]) 
Jv <- sort(Icv[(length(Icv)-298):length(Icv)]) # % 85%
Js <- Ics[(length(Ics)-99):length(Ics)] # % 5% 
```


```{r translate MATLAB code into R and explain the code}
s <- numeric(0)

thr <- 0.5
for (i in 1:100) {
  tt <- as.numeric(T[Js[i],])
  tm <- mean(tt)
  nm <- mean(as.numeric(N[Js[i],]))
  nmax <- max(as.numeric(N[Js[i],]))
  nmin <- min(as.numeric(N[Js[i],]))
  if (tm > nm) {
    Imax <- which(tt > nmax)
    sen <- length(Imax)/m
    if (sen > thr) {
      s <- c(s, as.numeric(Js[i]))
    }
  } else {
    Imin <- which(tt < nmin)
    sen <- length(Imin)/m
    if (sen > thr) {
      s <- c(s, as.numeric(Js[i]))
    }
  }
}

# Repeat for other thresholds
thr <- 0.7
for (i in 1:19) {
  tt <- as.numeric(T[Jm[i],])
  tm <- mean(tt)
  nm <- mean(as.numeric(N[Jm[i],]))
  nmax <- max(as.numeric(N[Jm[i],]))
  nmin <- min(as.numeric(N[Jm[i],]))
  if (tm > nm) {
    Imax <- which(tt > nmax)
    sen <- length(Imax)/m
    if (sen > thr) {
      s <- c(s, as.numeric(Jm[i]))
    }
  } else {
    Imin <- which(tt < nmin)
    sen <- length(Imin)/m
    if (sen > thr) {
      s <- c(s, as.numeric(Jm[i]))
    }
  }
}

thr <- 0.8
for (i in 1:299) {
  tt <- as.numeric(T[Jv[i],])
  tm <- mean(tt)
  nm <- mean(as.numeric(N[Jv[i],]))
  nmax <- max(as.numeric(N[Jv[i],]))
  nmin <- min(as.numeric(N[Jv[i],]))
  if (tm > nm) {
    Imax <- which(tt > nmax)
    sen <- length(Imax)/m
    if (sen > thr) {
      s <- c(s, as.numeric(Jv[i]))
    }
  } else {
    Imin <- which(tt < nmin)
    sen <- length(Imin)/m
    if (sen > thr) {
      s <- c(s, as.numeric(Jv[i]))
    }
  }
}

s0 <- sort(unique(s))
s0
```

Repeat the above process. But let T = T^2 and N = N^2
```{r}
T <- T^2
N <- N^2
m <- ncol(T) 
n <- ncol(N) 
Cm <- Cv <- Cs <- numeric(nrow(T))
for (i in 1:nrow(T)) {
  tm <- mean(as.numeric(T[i,]))
  nm <- mean(as.numeric(N[i,]))
  tv <- sd(as.numeric(T[i,]))
  nv <- sd(as.numeric(N[i,]))
  ts <- tv/tm
  ns <- nv/nm
  Cm[i] <- abs((tm-nm)/nm)
  Cv[i] <- abs((tv-nv)/nv)
  Cs[i] <- abs((ts-ns)/ns)
}
Icm <- order(Cm)
Icv <- order(Cv)
Ics <- order(Cs)
Jm <- sort(Icm[(length(Icm)-18):length(Icm)]) 
Jv <- sort(Icv[(length(Icv)-298):length(Icv)]) 
Js <- Ics[(length(Ics)-99):length(Ics)] 
s <- numeric(0)
thr <- 0.5
for (i in 1:100) {
  tt <- as.numeric(T[Js[i],])
  tm <- mean(tt)
  nm <- mean(as.numeric(N[Js[i],]))
  nmax <- max(as.numeric(N[Js[i],]))
  nmin <- min(as.numeric(N[Js[i],]))
  if (tm > nm) {
    Imax <- which(tt > nmax)
    sen <- length(Imax)/m
    if (sen > thr) {
      s <- c(s, as.numeric(Js[i]))
    }
  } else {
    Imin <- which(tt < nmin)
    sen <- length(Imin)/m
    if (sen > thr) {
      s <- c(s, as.numeric(Js[i]))
    }
  }
}
thr <- 0.7
for (i in 1:19) {
  tt <- as.numeric(T[Jm[i],])
  tm <- mean(tt)
  nm <- mean(as.numeric(N[Jm[i],]))
  nmax <- max(as.numeric(N[Jm[i],]))
  nmin <- min(as.numeric(N[Jm[i],]))
  if (tm > nm) {
    Imax <- which(tt > nmax)
    sen <- length(Imax)/m
    if (sen > thr) {
      s <- c(s, as.numeric(Jm[i]))
    }
  } else {
    Imin <- which(tt < nmin)
    sen <- length(Imin)/m
    if (sen > thr) {
      s <- c(s, as.numeric(Jm[i]))
    }
  }
}
thr <- 0.8
for (i in 1:299) {
  tt <- as.numeric(T[Jv[i],])
  tm <- mean(tt)
  nm <- mean(as.numeric(N[Jv[i],]))
  nmax <- max(as.numeric(N[Jv[i],]))
  nmin <- min(as.numeric(N[Jv[i],]))
  if (tm > nm) {
    Imax <- which(tt > nmax)
    sen <- length(Imax)/m
    if (sen > thr) {
      s <- c(s, as.numeric(Jv[i]))
    }
  } else {
    Imin <- which(tt < nmin)
    sen <- length(Imin)/m
    if (sen > thr) {
      s <- c(s, as.numeric(Jv[i]))
    }
  }
}
s <- sort(unique(s))
s
```

Then we find the same numbers in s and s0, which are the final gene IDs we want to get. The gene IDS are saved in "sfinal".
```{r}
sfinal <- s[s %in% s0]
sfinal
```

Through these codes, we can preliminarily screen out genes with significant differences between tumor people and normal people, and we will further screen these genes.


### MaxLogistic2.code

This function is to create logistic regression

The meaning of each parameter:
  X predictors
  ID response(normal is 0, and tumor is 1)
  n sampel size
  p the number of max groups
  beta the coefficient matrix
  I final predictor indexes
  q coefficients dimensions
```{r}
MaxLogistic2 <- function(X, ID, n, p, beta, I, q) {
  # Initialize XX matrix with negative infinities
  XX <- matrix(-Inf, nrow = n, ncol = p)

  # Index to manage column selection in X
  ii <- 1

  # Compute the linear combinations of beta and corresponding columns of X
  for (i in 1:p) {
    if (q[i] > 0) {
      subset_X <- cbind(X[, 1], X[, I[ii:(ii + q[i] - 1)] + 1])
      XX[, i] <- subset_X %*% beta[1:(q[i] + 1), i]
      ii <- ii + q[i]
    }
  }

  # Cap values at 10 to avoid numerical overflow
  XX[XX > 10] <- 10

  # Calculate the maximum value in each row
  Xmax <- apply(XX, 1, max)

  # Calculate probabilities using logistic function
  Pmax <- exp(Xmax) / (1 + exp(Xmax))

  # Create logical vectors for values <= 0.5 and > 0.5
  Pmax_lte_0_5 <- Pmax <= 0.5
  Pmax_gt_0_5 <- Pmax > 0.5

  # Calculate the sum for the objective function
  sg <- sum(Pmax_lte_0_5 * ID) + sum(Pmax_gt_0_5 * (1 - ID))

  return(sg)
}
```
sg means 

